%% This is an example first chapter.  You should put chapter/appendix that you
%% write into a separate file, and add a line \include{yourfilename} to
%% main.tex, where `yourfilename.tex' is the name of the chapter/appendix file.
%% You can process specific files by typing their names in at the 
%% \files=
%% prompt when you run the file main.tex through LaTeX.
\chapter{Introduction}

Graph mincut is a fundamental structure in graph theory with numerous applications. Let $G=(V,E)$ be an undirected unweighted connected graph on $n=|V|$ vertices and $m=|E|$ edges.
Two most common types of mincuts are global mincuts and pairwise mincuts. A set of edges with the least cardinality whose removal disconnects the graph is called a global mincut. For any pair of vertices $s,t\in V$, 
a set of edges with the least cardinality whose removal disconnects $t$ from $s$ is called a pairwise mincut for $s,t$ or simply a $(s,t)$-mincut. A more general notion is that of Steiner mincuts. For any given set $S\subseteq V$ of vertices, a set of edges with the least cardinality whose removal disconnects $S$ is called a Steiner mincut for $S$. It is easy to observe that the Steiner mincuts for $S=V$ are the global mincuts and for $S=\{s,t\}$ are $(s,t)$-mincuts.

While designing an algorithm for a graph problem, one usually assumes that the underlying graph is static. But, this assumption is unrealistic for most of the real-world graphs where vertices and/or edges do undergo change, though occasionally. Sensitivity Oracles can be used to efficiently make a query for a single update/delete operation. The data structure can be preprocessed offline.

% In the past, many elegant fault-tolerant algorithms have been designed for various classical problems, namely,  connectivity \cite{DBLP:conf/stoc/Chan02,DBLP:journals/algorithmica/FrigioniI00,DBLP:journals/siamcomp/ChanPR11,DBLP:conf/soda/DuanP17}, shortest-paths \cite{DBLP:conf/stoc/BernsteinK09,DBLP:journals/siamcomp/DemetrescuTCR08,DBLP:conf/stoc/ChechikC20}, graph spanners \cite{ChechikLPR10,DBLP:journals/tcs/BraunschvigCPS15}, SCC \cite{DBLP:journals/algorithmica/BaswanaCR19} , DFS tree \cite{DBLP:journals/siamcomp/BaswanaCC019}, and BFS structure \cite{DBLP:journals/talg/ParterP16,DBLP:journals/talg/ParterP18}. However, little is known about fault-tolerant data structures for various types of mincuts. The problem of fault-tolerant all-pairs mincuts aims at preprocessing a given graph to build a compact data structure so that the following query can be answered efficiently for any $s,t\in V$ and $(x,y)\in E$.\\

% \noindent
% FT-mincut$(u,v,x,y)$: Report the value of $(u,v)$-mincut in $G$ after the failure/removal of edge $(x,y)$, if exists, from $E$.\\

% \noindent
% % {\textsc{ft-mincut}}$(s,t,x,y)$: Report a $(s,t)$-mincut in $G$ after the failure of edge $(x,y)$, if exists, from $E$.\\
% {\textsc{ft-mincut}}$(s,t,x,y)$: Report a $(s,t)$-mincut in $G$ after the failure of edge $(x,y)$.\\



\section{Previous Results} 


There exists a classical ${\cal O}(n)$ size data structure that stores all-pairs mincuts \cite{GH61} known as Gomory-Hu tree. It is a tree on the vertex set $V$ that compactly stores a mincut between each pair of vertices. However, we cannot determine using a Gomory-Hu tree whether the failure of an edge will affect the $(s,t)$-mincut unless this edge belongs to the $(s,t)$-mincut present in the tree. We can get a fault-tolerant data structure by storing $m$ Gomory-Hu trees, one for each edge failure. The overall data structure occupies ${\cal O}(mn)$ space and takes ${\cal O}(1)$ time to report the value of $(s,t)$-mincut for any $s,t\in V$ upon failure of a given edge. A new $(s,t)$-mincut can itself be reported in ${\cal O}(n)$ time. 


For handling edge insertions, Baswana, Gupta and Knollman \cite{DBLP:conf/esa/BaswanaGK20} recently gave a data structure that can report the value of a $(s,t)$-mincut upon insertion of an edge. Moreover, a $(s,t)$-mincut incorporating the change can be reported in ${\cal O}(n)$ time.


Combining the above two results, a sensitivity oracle can be obtained that uses ${\cal O}(mn)$ space. However, the ${\cal O}(mn)$ space occupied by this data structure is far from the size of the graph. To the best of our knowledge, there exists no data structure for this problem which takes $o(mn)$ space and a non-trivial query time.



\section{Our Contribution} We present a space efficient sensitivity oracle for the all-pairs mincuts problem in an undirected unweighted multigraph. The data structure occupies ${\cal O}(n^2)$ space while achieving the optimal ${\cal O}(1)$ query time to report the value of $(s,t)$-mincut for any $s,t\in V$ upon deletion/insertion of any given edge. A resulting $(s,t)$-mincut incorporating the change can be reported in ${\cal O}(n)$ time.

In order to design our data structure, we present an efficient solution for a related problem of independent interest, called edge-containment query on a mincut defined as follows.\\

\noindent
{\textsc{edge-contained}}$(s,t,(x,y))$: Check if a given edge $(x,y)\in E$ belong to some $(s,t)$-mincut.\\

Using a data structure for this problem, we can answer a deletion query upon failure of edge $(x,y)$ by performing the corresponding edge-containment query. The value of $(s,t)$-mincut will reduce by unity if the edge-containment query evaluates to true. We can keep a Gomory-Hu tree of ${\cal O}(n)$ size as an auxiliary data structure to lookup the old value of $c_{s,t}$. The following fact allows us to do so.

\begin{fact}
\label{fact:(x,y)-lies-in-(s,t)-mincut}
The value of $(s,t)$-mincut decreases on deletion of an edge $(x,y)$ if and only if $(x,y)$ lies in \textit{some} $(s,t)$-mincut.
\end{fact}


We give conditional lower bounds for the following problems.

\begin{enumerate}
    \item Data structures that can handle dual-deletions (or dual-faults) for a $(s,t)$-mincut. In particular, we show that a data structure that can handle deletion (or failure) of two edges for a $(s,t)$-mincut even for fixed pair of vertices $s,t\in V$ requires ${\tilde \Omega}(n^2)$ space for non-trivial query time.
    \item Generalized flow tree for $2 \times 2$ mincuts, i.e. a data structure that can report the value of static $(\{s,u\},\{v,t\})$-mincut for given vertices $s,t,u,v \in V$. We show that even if we fix $s,t \in V$, the data structure must require ${\tilde \Omega}(n^2)$ space for non-trivial query time.
\end{enumerate}


\section{Related Work} 
A related problem is that of maintaining mincuts in a dynamic environment. Until recently, most of the work on this problem has been limited to global mincuts. Thorup \cite{DBLP:journals/combinatorica/Thorup07} gave a Monte-Carlo algorithm for maintaining a global mincut of polylogarithmic size with ${\tilde {\cal O}}(\sqrt{n})$ update time.  He also showed how to maintain a global mincut of arbitrary size with $1+o(1)$-approximation within the same time-bound. Goranci, Henzinger and Thorup \cite{DBLP:journals/talg/GoranciHT18} gave a deterministic incremental algorithm for maintaining a global mincut with amortized ${\tilde{\cal O}}(1)$ update time and ${\cal O}(1)$ query time. Hartmann and Wagner \cite{DBLP:conf/isaac/HartmannW12} designed a fully dynamic algorithm for maintaining all-pairs mincuts  which provided significant speedup in many real-world graphs, however, its worst-case asymptotic time complexity is not better than the best static algorithm for an all-pairs mincut tree. Recently, there is a fully-dynamic algorithm \cite{DBLP:journals/corr/abs-2005-02368} that approximates all-pairs mincuts up to a nearly logarithmic factor in ${\tilde{\cal O}}(n^{2/3} )$ amortized time against an oblivious adversary, and ${\tilde{\cal O}}(m^{3/4} )$ time against an adaptive adversary. To the best of our knowledge, there exists no non-trivial dynamic algorithm for all-pairs \textit{exact} mincut. We feel that our insights in this paper may be helpful in this problem.

\noindent


% \section{Overview of our results} 

% Dinitz and Vainshtein \cite{DBLP:journals/siamcomp/DinitzV00}
% presented a novel data structure called {\em connectivity carcass} that stores all Steiner mincuts for a given Steiner set $S\subseteq V$ in ${\cal O}(\min(m,nc_S))$ space, where $c_S$ is the value of Steiner mincut. %We observe that this data structure can be used easily to design a fault-tolerant data structure for all-pairs mincuts that occupies $O(mn)$-space. 
% Katz, Katz, Korman and Peleg \cite{DBLP:journals/siamcomp/KatzKKP04} presented a data structure of ${\cal O}(n)$ size for labeling scheme of all-pairs mincuts. This structure hierarchically partitions the vertices based on their connectivity in the form of a rooted tree. In this tree, each leaf node is a vertex in set $V$ and each internal node $\nu$ stores the Steiner mincut value of the set $S(\nu)$ of leaf nodes in the subtree rooted at $\nu$. %stores all-pairs mincut value.
% % Apparently, both these data structures seem to be designed for a static graph. 
% We observe that if each internal node $\nu$ of the hierarchy tree
% is augmented with the connectivity carcass of $S(\nu)$, we get a data structure for the edge-containment query. This data structure occupies ${\cal O}(mn)$ space. An edge-containment query can be answered using the connectivity carcass at the Lowest Common Ancestor (LCA) of the given pair of vertices.
% %following, more generic, fault-tolerant query :

% As we move down the hierarchy tree, the size of Steiner set associated with the internal node reduces. So, to make the data structure more compact, a possible approach is to associate a smaller graph $G_\nu$ for each internal node $\nu$ that is {\em small enough} to improve the overall space-bound, yet {\em large enough} to retain the internal connectivity of set $S(\nu)$. 
% However, such a compact graph cannot directly answer the edge-containment query as it does not even contain the information about all edges in $G$. A possible way to overcome this challenge is to transform any edge-containment query in graph $G$ to an equivalent query in graph $G_{\nu}$. In this paper, we show that not only such a transformation exists, but it can also be computed efficiently. We model the query transformation as a multi-step procedure. The following result captures a single step of this procedure.

% % The key observation that allows us to do query transformation is as follows. 
% Given an undirected graph $G=(V,E)$ and a Steiner set $S\subseteq V$, let $S'\subset S$ be any maximal set with connectivity strictly greater than that of $S$. We can build a quotient graph $G_{S'}=(V_{S'},E_{S'})$ such that $S' \subset V_{S'}$ with the following property.

% {\em
% For any two vertices $s,t \in S'$ and any set of edges $E_y$ incident on vertex $y$ in $G$, there exists a set of edges $E_{y'}$ incident on a vertex $y'$ in $G_{S'}$ such that $E_y$ lies in a $(s,t)$-mincut in $G$ if and only if $E_{y'}$ lies in a $(s,t)$-mincut in $G_{S'}$. 
% }

% We build graph $G_\mu$ associated with each internal node $\mu$ of the hierarchy tree as follows. For the root node $r$, $G_r = G$. For any other internal node $\mu$, we use the above result to build the graph $G_{\mu}$ from $G_{\mu'}$, where $\mu'$ is the parent of $\mu$. 
% Our data-structure is the hierarchy tree where each internal node $\mu$ is augmented with the connectivity carcass for $G_\mu$ and the Steiner set $S(\mu)$. A high-level description of our query algorithm is as follows. 
% % We move from the root node to the LCA of $s$ and $t$.
% We traverse the path from the root node to the LCA of $s$ and $t$.
% % We keep transforming an edge-containment query for each edge in this path. 
% We keep transforming the edge-containment query for each edge in this path.
% At the LCA of $s$ and $t$, we stop and perform the query using the connectivity carcass 
% % augmented at this node
% stored at this node.
% Following a rigorous analysis, we show that this data structure takes only ${\cal O}(m)$ space and can answer any edge-containment query in ${\cal O}(\min(m,nc_{s,t}))$ time.

% \section{Organization of the paper:}

% In addition to the basic notations, Section \ref{sec:prelimiaries} presents compact representation for various mincuts. Section \ref{sec:query-transformation} gives insights into $3$-vertex mincuts that form the foundation for transforming an edge-containment query in original graph to a compact graph. We give the construction of compact graph for query transformation in Section \ref{sec:compact-graph-section}.
% Using this compact graph as a building block, we present the data structures for edge-containment query in Section \ref{sec:final-ds}.